USA pollster

### 2024-11-04 paper/paper.qmd
Refining appendix A and B
Appendix A:
1. Added a simulation for a Patriot Polling methodology for Wisconsin poll.
2. Elaboration on sampling bias using tables
3. Specified population and sampling frame

Appendix B:
1. Completed budgeting for idealised methodology
2. Included google form link under pilot section

Note: I am using Ipums census data which is too big for github.

library(tidyverse)
library(palmerpenguins)
library(knitr)
```

@@

## Outcome variables

@@

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

@@

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

## Methodology
From 12 to 14 October 2024, Patriot Polling has conducted phone surveys in the state of Wisconsin of the United States of America, which borders Minnesota, Iowa, Michegan, and Illinois. Their target population are voters in Wisconsin. For sample of landlines which are characterised by households, an equal number of phone numbers are randomly selected across every landline block in Wisconsin using Random Digit Dialing (RDD). As for the sample of mobile numbers, which are characterised by an individual with access to a digital mobile device, the sample was purchased from a consumer contact number database. For this poll, a total of 803 respondents has completed the survey either from contact through the landline or their personal contact number. [@patriotPollingMethodology]

@patriotPollingMethodology conducted polling in such a way where a randomly contacted respondent would hear pre-recorded voice messages and users are able to interact with the automated phone system. This system is called Interactive voice response (IVR) which can handle outbound calls more systematically since the same questions would be asked sequentially in the survey. There were only 2 questions that were asked. Both questions begin with 'How will you vote for president?'. This is followed by the candidates name as it appears in the voting ballot.

@@

Lastly, the purchasing of phone numbers from a consumer database would introduce of selection bias. Individuals from middle to higher socio-economic status would have greater purchasing power than their lower income counterparts, which means individuals from lower socio-economic status may not have phones. In addition, elderlies within Wisconsin may not necessarily have mobile phones due to the lack of skills to use such technologies. Also, individuals would've opted for do-not-call registries, opted out of marketing databases or simply have prepaid phones which do not link to their personal information. This could mean an under representation of individuals from the said groups.

```{r senate they voted for}
#| label: tbl-senate
#| include: false
#| echo: false
#| layout-ncol: 2
#| fig-cap: "Presidental Candidates Votes"
#| fig-subcap: ["Percentage of votes for senate candidates in Wisconsin", "Percentage of votes for presidential candidates in Wisconsin"]

presidential_voter_table <- tibble(
  'Presidential Candidate'= c('Donald Trump (R)', 'Kamala Harris (D)'),
  'Registered Voters'= c('50%', '49%'),
)
senate_voter_table <- tibble(
  'Senate Candidate'= c('Tammy Baldwin (R)', 'Eric Hovde (D)'),
  'Registered Voters'= c('50%', '49%'),
)

kable(presidential_voter_table) 
kable(senate_voter_table)
```

@@

## Idealised methodology: Pilot 
The purpose of a pilot testing is to ensure a well-designed election forecasting survey. Due to the nature of the general election characterised by the votes of the electoral college, random sampling has to be performed across all 50 states of America. \$5,000 should go towards a pilot test obtaining 200 respondents. This pilot test would target battleground states such as Wisconsin as well as Maine and Nebraska to test specific issues around a split vote between the two presidential candidate. Begin by using stratified sampling that divides each state population into strata by various socio demographic factors (e.g. gender, highest educational attainment, race, income), then conduct systematic random sampling within each strata. An optional \$5 digital or mailed voucher can be used to incentivise respondents to complete the survey.

@@

In order to expand the sampling frame to include voters who may not prefer to do a phone surveys, we can conduct a Google forms survey as well which is free of charge. The format of the form should mimic that of the RDD random sampling method explained in the paragraph before this.

@@

## Idealised methodology: Large scale sampling
By refining our phone surveys and google form surveys, we can expand the surveys to all states in America. For the states of Nebraska and Maine, we will be performing both analysis for individual congressional districts and statewide majority votes to determine the number of estimated electoral votes as described in @sec-NebraskaMaine. Particularly for all battleground states as seen from the past 2020 elections like Arizona, Florida and Maine's second congressional district, we will be oversampling to ensure that the preferences of sampled voters are better representive of the overall voter population in their particular state or congressional district, leading to reduced sampling error and more precise forecasting of the presidential elections.

@@

Ultimately, the algorithm to forecast the presidential winner involves obtaining the majority vote based on samples obtained within each state (and districts in Nebraska and Maine). Then based on the  number of electors assigned to that region, we will sum up the number of electors affiliated to the presidential candidates. This sum divided by the 538 total electors would be the predicted probability that a presidential candidate would win in the upcoming election.

@@

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

>>>>

library(tidyverse)
library(palmerpenguins)
library(knitr)
data <- read_csv("../data/02-analysis_data/analysis_data.csv")
```

@@

## Predictor Variables

@@

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
data |>
  select(poll_id, pollster_id, candidate_name, winner)
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

@@

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
```

```{r}
raw_pollster_data <- read_csv('../data/01-raw_data/raw_pollster_data.csv')

raw_pollster_data |>
  colnames()

raw_pollster_data |>
  select(party, candidate_name) |>
  unique() |>
  arrange(party, candidate_name)

raw_pollster_data |>
  select(party, candidate_name) |>
  unique() |>
  filter(grepl('Donald', candidate_name))

# remove trailing time stamp
raw_pollster_data$created_at <- sapply(strsplit(raw_pollster_data$created_at, '\\s+'), '[', 1)

time_var <- c('start_date', 'end_date', 'created_at')

# get day, month, year in character data type
raw_pollster_data$month <- sapply(strsplit(raw_pollster_data$created_at, '/'), '[', 1)
raw_pollster_data$day <- sapply(strsplit(raw_pollster_data$created_at, '/'), '[', 2)
raw_pollster_data$year <- sapply(strsplit(raw_pollster_data$created_at, '/'), '[', 3)

# observe how year is not cleaned
unique(raw_pollster_data$year)

# correct different formats for year
raw_pollster_data$year <- case_when(
  raw_pollster_data$year %in% c('24', '2024') ~ 2024,
  raw_pollster_data$year %in% c('23', '2023') ~ 2023,
  raw_pollster_data$year %in% c('22', '2022') ~ 2022,
  raw_pollster_data$year %in% c('21', '2021') ~ 2021,
  )

# check that year is cleaned
unique(raw_pollster_data$year)

raw_pollster_data <- raw_pollster_data |>
  mutate(day = ifelse(day < '10', paste0('0', day), day),
         month = ifelse(month < '10', paste0('0', month), month),
         created_at = as.Date(paste(year, month, day, sep='-')))

# get day, month, year in character data type
raw_pollster_data$month <- sapply(strsplit(raw_pollster_data$start_date, '/'), '[', 1)
raw_pollster_data$day <- sapply(strsplit(raw_pollster_data$start_date, '/'), '[', 2)
raw_pollster_data$year <- sapply(strsplit(raw_pollster_data$start_date, '/'), '[', 3)

# observe how year is not cleaned
unique(raw_pollster_data$year)

# correct different formats for year
raw_pollster_data$year <- case_when(
  raw_pollster_data$year %in% c('24', '2024') ~ 2024,
  raw_pollster_data$year %in% c('23', '2023') ~ 2023,
  raw_pollster_data$year %in% c('22', '2022') ~ 2022,
  raw_pollster_data$year %in% c('21', '2021') ~ 2021,
  )

# check that year is cleaned
unique(raw_pollster_data$year)

raw_pollster_data <- raw_pollster_data |>
  mutate(day = ifelse(day < '10', paste0('0', day), day),
         month = ifelse(month < '10', paste0('0', month), month),
         start_date = as.Date(paste(year, month, day, sep='-')))

# get day, month, year in character data type
raw_pollster_data$month <- sapply(strsplit(raw_pollster_data$end_date, '/'), '[', 1)
raw_pollster_data$day <- sapply(strsplit(raw_pollster_data$end_date, '/'), '[', 2)
raw_pollster_data$year <- sapply(strsplit(raw_pollster_data$end_date, '/'), '[', 3)

# observe how year is not cleaned
unique(raw_pollster_data$year)

# correct different formats for year
raw_pollster_data$year <- case_when(
  raw_pollster_data$year %in% c('24', '2024') ~ 2024,
  raw_pollster_data$year %in% c('23', '2023') ~ 2023,
  raw_pollster_data$year %in% c('22', '2022') ~ 2022,
  raw_pollster_data$year %in% c('21', '2021') ~ 2021,
  )

# check that year is cleaned
unique(raw_pollster_data$year)

raw_pollster_data <- raw_pollster_data |>
  mutate(day = ifelse(day < '10', paste0('0', day), day),
         month = ifelse(month < '10', paste0('0', month), month),
         end_date = as.Date(paste(year, month, day, sep='-')))

raw_pollster_data |>
  select(start_date, end_date, created_at) |>
  summarise(min(start_date))

sum(is.na(raw_pollster_data$partisan))

raw_pollster_data |>
  filter(!is.na(state))

raw_pollster_data |>
  select(office_type) |>
  unique()

raw_pollster_data |>
  summarise(min(pollscore, na.rm=T), max(pollscore, na.rm=T))

raw_pollster_data |>
  summarise(min(transparency_score, na.rm=T), max(transparency_score, na.rm=T))

raw_pollster_data |>
  summarise(min(numeric_grade, na.rm=T), max(numeric_grade, na.rm=T))

raw_pollster_data |>
  summarise(min(start_date, na.rm=T), max(start_date, na.rm=T),
            min(end_date, na.rm=T), max(end_date, na.rm=T))
```

@@

## Methodology: Overview {#sec-patriot_polling_overview}
From 12 to 14 October 2024, Patriot Polling has conducted phone surveys in the state of Wisconsin of the United States of America, which borders Minnesota, Iowa, Michegan, and Illinois. Their target population are voters in Wisconsin. For sample of landlines which are characterised by households, an equal number of phone numbers are randomly selected across every landline block in Wisconsin using Random Digit Dialing (RDD). As for the sample of mobile numbers, which are characterised by an individual with access to a digital mobile device, the sample was purchased from a consumer contact number database. For this poll, a total of 803 respondents has completed the survey either from contact through the landline or their personal contact number. [@patriotPollingMethodology]

@patriotPollingMethodology conducted polling in such a way where a randomly contacted respondent would hear pre-recorded voice messages and users are able to interact with the automated phone system. This system is called Interactive voice response (IVR) which can handle outbound calls more systematically since the same questions would be asked sequentially in the survey. There were 2 questions provided by Patriot Polling that were asked to respondents. Both questions begin with 'How will you vote for president?'. This is followed by the candidates name as it appears in the voting ballot. There was no indication whether this was all the questions that were asked but there were tables provided by the pollster showing the breakdown by sex, education, preferred political affiliation and income. Therefore, it would be safe to assume that there were questions related to the socio-demographic characteristics of the respondents even though we don't know the specific questions asked.

## Methodology: Population
The targeted population of this poll was not explicitly stated. While the pollster did conduct phone surveys, they did not mention any inclusion or exclusion criteria. Furthermore, the use of IVR during phone surveys and lack of information provided regarding the questions asked as discussed in @sec-patriot_polling_overview, it will be difficult to guarantee that the polls are targeting all eligible voters in Wisconsin. Given that nature of the polling article focuses on Wisconsin, it would be safe to assume that the targeted population would be eligible voters above 18 years old in Wisconsin.

## Methodology: Sampling Frame
Since phone surveys were conducted on both landlines and smartphones, the sampling frame of this poll would be eligible voters above 18 years old in Wisconsin with a working landline or possess a phone with a working contact number.

@@

Overall, the biggest weakness in the methodology would be the purchasing of phone numbers from a consumer database as it introduces selection bias. Individuals from middle to higher socio-economic status would have greater purchasing power than their lower income counterparts, which means individuals from lower socio-economic status may not have phones. In addition, elderlies within Wisconsin may not necessarily have mobile phones due to the lack of skills to use such technologies. Also, individuals would've opted for do-not-call registries, opted out of marketing databases or simply have prepaid phones which do not link to their personal information. This could mean an under representation of individuals from the said groups.

## Simulation exploring weaknesses in Patriot's Polling's methodology
To illustrate the biggest concern raised in @sec-WisconsinWeaknesses, we will be simulating a Patriot's Polling's methodology by using US Census data to obtain random our pool of samples that are representive of the population of Wisconsin with information about their age, race and county. Since the raw data provided by IPUM [@USACensusData] encoded states and counties, we combined the raw data with a supplementary excel file provided by IPUM to obtain the actual county name. We will be using the variable $CISMRTPHN$ from the census data which represents whether the particular respondent of the census data has a smartphone. This variable will be used to mirror a Patriot's Polling methodology of obtaining personal contact number by purchasing from an online database. We used this approach as only Wisconsin respondents with smartphones can possibly be included in a smartphone contact number database.

For the sake of discussion, we will be obtaining 2 samples. **Sample A will include respondents above the voting age of 18 in Wisconsin regardless of whether they possess a smartphone while sample B will only include respondents above the voting age of 18 in Wisconsin who have smartphones.** Sample B will mimic purchasing phone numbers from a database.

@tbl-race_correct_representation shows us the proportion of white, black and asian races from sample A while @tbl-race_wrong_representation shows us that from sample B. We notice that black individuals from Wisconsin are underrepresented (8.7% vs 9.4%). According to @BlackPovertyWisconsin, black poverty rates in Wisconsin are 2.5 times higher than the overall Wisconsin poverty rate. This supports the finding that if we were to include only individuals with smartphones, certain racial groups might be underrepresented. Likewise, we are able to observe Asians being overrepresented (4.4% vs 3.8%). @AsianWealthUS exerts that for various educational attainment levels, Asians are earning more than individuals who belong to white, black or hispanic households. It is unsurprising that Asians are being overrepresented as higher wealth among the Asian population in Wisconsin would translate to higher purchasing power to purchase smartphones. Asians would naturally be more likely to be included in a sample where only smartphone users are considered, by extension, Asians would be more likely to be included in a sample obtained from a consumer database containing user contact numbers.
```{r}
#| echo: false
#| eval: true
#| label: tbl-race_correct_representation
#| tbl-cap: "proportion of races from non-smartphone-discriminating sample"
#| warning: false
census_data <- read_csv("../data/00-simulated_data/simulated_appendix_data.csv")

set.seed(304)
smartphone_only_sample <- census_data |>
  filter(has_smartphone == 1) |>
  sample_n(1000, replace=FALSE, weight = weight)

all_sampled <- census_data |>
  sample_n(1000, replace=FALSE, weight = weight)

race_table_full_sample <- all_sampled |> 
  group_by(race) |>
  count() |>
  ungroup() |>
  mutate(prop = paste0(round(n/sum(n) * 100,1),'%')) |>
  select(race, count = n, proportion = prop)

kable(race_table_full_sample)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-race_wrong_representation
#| tbl-cap: "proportion of races from smartphone-only sample"
#| warning: false

# asians overrepresented, blacks underrepresented
race_table_smartphone_sample <- smartphone_only_sample |> 
  group_by(race) |>
  count() |>
  ungroup() |>
  mutate(prop = paste0(round(n/sum(n) * 100,1),'%')) |>
  select(race, count = n, proportion_chosen = prop)

kable(race_table_smartphone_sample)
```

This becomes an issue because underrepresentation of black communities and overrepresentation of asian communities in Wisconsin would not reflect the actual voting preferences of a particular racial demographic, resulting in sampling bias, potentially misleading people who interpret a Patriot Polling's poll on Wisconsin or even their written articles.

Next, @tbl-age-differerences shows us the the percentage difference in individuals chosen categorised by age groups from sample A and B. We are immediately alerted to a 20% decrease in the number of elderlies chosen from sampling method B compared to sampling method A. As of 2024, elderlies above the age of 65 are either from the baby boomer generation or silent generation, who are known to experience difficulty when using smartphones. @SmartphoneByAge shows us that 97% of Americans between 18 to 49 have smartphones, 89% for that of age groups 50 to 64, and a lower 76% of the American elderlies above 65 year old who own smartphones. Therefore, by proceeding with sample B (parallel to purchasing phone numbers from consumer database) for which the inclusion critera selects individuals who own smartphones, we might be discounting the votes of elderly population, resulting in selection bias.

{{< pagebreak >}}

```{r}
#| echo: false
#| eval: true
#| label: tbl-age-differerences
#| tbl-cap: "percentage difference in individuals chosen from sample A and B by age groups"
#| warning: false

################
#      age 
#
# results: median age is lower, and more elderlies above 65 years old
#          are underrepresented when obtaining phone numbers straight from 
#          database
################ 
age_bin_count_full <- all_sampled |>
  count(age_bin)

age_table <- smartphone_only_sample |>
  count(age_bin) |>
  inner_join(age_bin_count_full,
             by = c('age_bin'='age_bin')) |>
  mutate(diff = n.x - n.y,
         perc_diff_1 = round(diff/n.y * 100, 0),
         perc_diff_2 = case_when(
           diff < 0 ~ paste0(perc_diff_1, '%'),
           .default = paste0('+', perc_diff_1, '%'))) |>
  arrange(desc(abs(perc_diff_1))) |>
  select(`age bin` = age_bin, 
         `sample A` = n.y,
         `sample B` = n.x, 
         `perc diff` = perc_diff_2)

kable(age_table)
```

{{< pagebreak >}}

@@

## Idealised methodology: Pilot 
The purpose of a pilot testing is to ensure a well-designed election forecasting survey. Due to the nature of the general election characterised by the votes of the electoral college, random sampling has to be performed across all 50 states of America. \$5,000 should go towards a pilot test obtaining 200 respondents per state. This pilot test would target battleground states such as Wisconsin as well as Maine and Nebraska to test specific issues around a split vote between the two presidential candidate. Begin by using stratified sampling that divides each state population into strata by various socio demographic factors (e.g. gender, highest educational attainment, race, income), then conduct systematic random sampling within each strata. An optional \$5 digital or mailed voucher can be used to incentivise respondents to complete the survey.

@@

In order to expand the sampling frame to include voters who may not prefer to do a phone surveys, we can conduct a Google forms survey as well which is free of charge. The format of the form should mimic that of the RDD random sampling method. Refer to [this google form](https://forms.gle/aMUwiEs9yNwpT4tq8) for an example of the survey questions aksed.

@@

## Idealised methodology: Large scale sampling
By refining our phone surveys and google form surveys, we can expand the surveys to all states in America. For this we budget out \$80,000 to obtain between 2000 to 5000 samples across each state, depending on the state's population. For the states of Nebraska and Maine, we will be performing both analysis for individual congressional districts and statewide majority votes to determine the number of estimated electoral votes as described in @sec-NebraskaMaine. Particularly for all battleground states as seen from the past 2020 elections like Arizona, Florida and Maine's second congressional district, we will be oversampling to ensure that the preferences of sampled voters are better representive of the overall voter population in their particular state or congressional district, leading to reduced sampling error and more precise forecasting of the presidential elections.

@@

Ultimately, the algorithm to forecast the presidential winner involves obtaining the majority vote based on samples obtained within each state (and districts in Nebraska and Maine). Then based on the  number of electors assigned to that region, we will sum up the number of electors affiliated to the presidential candidates. This sum divided by the 538 total electors would be the predicted probability that a presidential candidate would win in the upcoming election. \$10,000 will be spent on the cost of surveyors. The remaining \$5,000 from our budget will be used as contigency funds.

@@

{{< pagebreak >}}

@@

```
